### 2025-04-26 paper2/Paper.tex
update-paper
Bayesian-framework-determined

% Effort, Output and Gap
We assume two players, $i$ and $j$, compete for a prize $\theta>0$ in a contest. 
Winner gets the prize and loser gets nothing. 
The contest starts at time zero. 
At every time $t\ge0$, the representative player $i$ chooses an effort level $q_{i,t}$ and burdens a quadratic cost $C_i(q_{i,t}) = c_i q_{i,t}^2/2$, with a lower $c_i$ corresponding to higher ability. 
Denoted by $y_t$ the \textit{output gap} of player $i$ and $j$ at time $t$, driven by
\begin{equation}
	dy_t = (q_{i,t}-q_{j,t})dt + \sigma dW_t
\end{equation}
where $W_t$ is a Brownian motion and $\sigma>0$ measures the innovation risk.

We assume that the contest is equipped with a submission system that allows participants to upload their algorithms at any time and receive immediate feedback. 
For simplicity, we further assume that agents submit their intermediate results whenever they make progress. 
This setup enables the contest organizers to monitor all players’ progress $x_{i,t}$ and $x_{j,t}$ in real time.
This true output level, evaluated by the system, is only known by the game designer but not the two players.

% Signal
At any time $t>0$, the contest designer emits a \textit{public} signal of the real output gap $y_t$. 
The signal is ambiguous and the game holder controls the ambiguity. 
The dynamic of signal is  
\begin{equation}\label{signal}
	dZ_{t} = y_{t}dt + \frac{dB_{t}}{\sqrt{\lambda}} 
\end{equation}
where $B_{t}$ is standard Brownian motion independent with $(W_{i,t})$ and $(W_{j,t})$, and the parameter $\lambda$ is set by the game holder to control the precision of signal. 
The larger the $\lambda$, the more accurate the signal would be.

% Bayesian Player
The information set of both players at time $t \ge 0$ is  $I_{t} \equiv \{Z_{s} : 0\le s \le t\}$. 
Player $i$ estimates the unknown output gap $y_t$ based on the information set $I_t$. 
Let $\tilde{y}_t \equiv E(y_{t}|I_t)$ be the estimated output gap and $S_t \equiv E[(\tilde{y}_{t}-y_{t})^2|I_t]$ be the estimation variance. 
According to Chapter 1.2 of \citet{Bensoussan1992Control}, \textit{Kalman-Bucy filter} gives the dynamics of $\tilde{y}_{t}$ and $S_{t}$, 
\begin{align}
	d\tilde{y}_{t} &= (q_{i,t}-q_{j,t})dt + \lambda S_{t}(dZ_{t}-\tilde{y}_{t}dt) 
	\label{filtered-x}\\
	\frac{dS_{t}}{dt} &= \sigma^2 - \lambda S_{t}^2
	\label{filtered-S}
\end{align} 
Hence, the conditional distribution $y_{t}|I_t\sim\mathcal{N}(\tilde{y}_{t},S_{t}|I_t)$ is fully captured by the mean $\tilde{y}_{t}$ and variance $S_{t}$. 
If $\lambda = 0$, we have $S_t = S_0+\sigma^2t$, i.e., the estimation variance is increasing in time linearly. 
If $\lambda > 0$, the solution of (\ref{filtered-S}) is
\begin{equation}\label{S-evolution} 
	S_t = 
	\begin{cases}
        \bar{S} \cdot \tanh\left\{t\cdot\sigma\sqrt{\lambda} +  \tanh^{-1}\left(S_0\big/\bar{S}\right)\right\} & \text{if } S_0 < \bar{S} \\
        \bar{S} & \text{if } S_0 = \bar{S} \\
        \bar{S} \cdot \coth\left\{t\cdot\sigma\sqrt{\lambda} +  \coth^{-1}\left(S_0\big/\bar{S}\right)\right\} & \text{if } S_0 > \bar{S} 
	\end{cases}
\end{equation}
Specifically, $\bar{S} = \sigma / \sqrt{\lambda}$ when $\lambda>0$ and $\bar{S} = \infty$ when $\lambda = 0$. 
Please refer to Appendix~\ref{app-S-equ} for the derivations. 
Figure~\ref{fig-S-evol} shows the evolution of $S_t$ in time: estimation variance $S_t$ converges to \textit{steady state} $\bar{S}$ as time goes by regardless of the starting estimation variance. 
For simplicity, we henceforth assume that $S_0 = \bar{S}$, hence $S_t\equiv \bar{S}$.

\begin{figure}[!ht]
	\centering
	\includegraphics[scale=0.6]{figure_S_evolve.pdf}
	\caption{The evolution of $S_t$ in time $t$, given that $\lambda = 1$ and $\sigma = 1$.} \label{fig-S-evol}
\end{figure}

% Contest Deadline
Following \cite{ryvkin2022fight}, let's consider a dynamic contest with a fixed deadline. 
Suppose the contest is terminated when time $t=T>0$. 
Since the steady state estimation variance $\bar{S}$ is fixed as displayed above, the state of the game is fully characterized by a tuple $(\tilde{y}_t, t)$. 
At any time $0\le t<T$, player $i$ optimizes her effort level $q_{i,\tau}$ in the remaining contest period $\tau\in[t, T)$ according to the following optimization problem, 
\begin{equation}\label{v-def}
	V^i(\tilde{y}_{t}, t ; q_{j,t},\Theta_i) = \max_{\{q_{i,\tau}\}^T_{\tau=t}} 
	\mathbb{E}\left( \theta\cdot1_{\tilde{y}_T>0} - \int^T_tC_i(q_{i,\tau})d\tau \bigg|I_t\right) 
\end{equation}
where $\Theta_i \equiv\{\theta, \lambda, \sigma, c_i\}$, subject to constraints (\ref{filtered-x}), (\ref{filtered-S}) and $q_{i,\tau}\ge0$ for all $\tau\in[t,T)$.
The optimization problem for player $j$ is just symmetric to that of player $i$ as $V^j(\tilde{y}_t, t) = V^i(-\tilde{y}_t, t)$. 
The corresponding Hamilton-Jacobi-Bellman (HJB) equation for player $i$ is 
\begin{equation*}
0 = \max_{q_{i,t}\ge0}\left[-\frac{c_iq_i^2}{2} + V^i_{y}\cdot\left(q_{i,t}-q_{j,t}\right)+V^i_t + \frac{V^i_{yy}}{2}\lambda \bar{S}^2\right]
\end{equation*}
By definition, we have $\lambda \bar{S}^2 = \sigma^2$. 
Under the assumption of inner solution, we plug into the first order conditions $q_{i,t} = V^i_y/c_i$ and $q_{j,t} = -V^j_y/c_j$, we have the system of equations
\begin{equation*}
\begin{aligned}
\frac{1}{2c_i}(V^i_y)^2 + \frac{1}{c_j}V^i_yV^j_y + V^i_t + V^i_{yy}\frac{\sigma^2}{2} = 0\\
\frac{1}{2c_j}(V^j_y)^2 + \frac{1}{c_i}V^j_yV^i_y + V^j_t + V^j_{yy}\frac{\sigma^2}{2} = 0
\end{aligned}
\end{equation*}
subject to boundary conditions $V^i(-\infty, t) = 0$, $V^i(+\infty, t) = \theta$, $V^j(-\infty, t) = \theta$, $V^j(+\infty, t)=0$, $V^i(\tilde{y}_T, T) = \theta \cdot 1_{\tilde{y}_T > 0}$ and $V^j(\tilde{y}_T, T) = \theta \cdot 1_{\tilde{y}_T < 0}$.

@@

The likelihood function of any realization of this point process $\{t^i_k\}_{k=1}^{N_i}$ is given by 
\begin{equation}\label{eq-ihpp-prob}
p\left(\{t^i_k\}_{k=1}^{N_i} | \tau_i\right) = \exp\left\{-\int_{s\in\mathcal{T}}\tau_i(s)ds\right\}\prod_{k=1}^{N_i}\tau_i(t^i_k)
\end{equation}

@@

Based on the following assumptions, we will next generate the submission data for two players in a data analysis competition: 
\begin{itemize}
	\item Total uncertainty: $\sigma = 5$
	\item Capacities: $c_i = 1.5$ and $c_j = 2$
	\item Contest duration: from 2025-01-01 to 2025-03-31
	\item Starting point $\tilde{y}_0=0$
\end{itemize}

Simulation algorithm:

\begin{algorithm}
\caption{Synthetic Data Simulation}
\begin{algorithmic}
\STATE \textbf{Input:} 
	$\{\hat{t}^i_k\}_{k=1}^{N_i}$, $c_i$, $c_j$, $\sigma$, $r$.
\STATE Initialize $\tilde{y}_t = 0$, $\hat{y}_t = 0$
\FOR{$t =$ 2025-01-01 to 2025-03-31}
    \STATE $m_i(\tilde{y}_t, t), m_j(\tilde{y}_t, t)$  $\gets$ (\ref{eq-equilibrium-effort})
    \STATE $\tilde{y}_{t+\Delta t}$ $\gets$ (\ref{eq-fintered-y-update})
\ENDFOR
\STATE \textbf{Output:} $(\tilde{y}_t)$, $(\hat{y}_t)$
\end{algorithmic}
\end{algorithm}

@@

\newpage
\bibliography{_Literatures.bib}

@@

\section{Solve $S_t$ in Equation (\ref{filtered-S})}\label{app-S-equ}

If $S$ is in steady state $dS/dt = 0$ $\Leftrightarrow$ $S = \bar{S} \equiv \sigma/\sqrt{\lambda}$. 
If $S$ is not in steady state, i.e. $S\not=\bar{S}$, we first isolate the two variables and get 
\begin{equation*}
	\frac{dS}{\sigma-\lambda S^2} = dt
\end{equation*}
Then, we take the integral on both sides
\begin{equation*}
    t = \int \frac{dS}{\sigma-\lambda S^2} = \frac{1}{\sigma\sqrt{\lambda}} \int \frac{dS\sqrt{\lambda}/\sigma}{1 - (S\sqrt{\lambda}/\sigma)^2} \equiv \frac{1}{\sigma\sqrt{\lambda}}\int \frac{du}{1-u^2} 
\end{equation*}
where $u = S\sqrt{\lambda}/\sigma = S/\bar{S}$. Hence, 
\begin{equation*}
    \sigma\sqrt{\lambda}\cdot t = \begin{cases}
        \tanh^{-1}(u)-K_1, & \text{if } |u|<1\\
        \coth^{-1}(u)-K_2, & \text{if } |u|>1
    \end{cases} = \begin{cases}
        \tanh^{-1}(S/\bar{S})-K_1, & \text{if } S<\bar{S}\\
        \coth^{-1}(S/\bar{S})-K_2, & \text{if } S>\bar{S} 
\end{cases}
\end{equation*}
Thus, we conclude the non-steady state case that 
\begin{equation*}
	S = \begin{cases}
		\bar{S}\cdot\tanh(\sigma\sqrt{\lambda}\cdot t + K_1), & \text{if } S<\bar{S}\\
		\bar{S}\cdot\coth(\sigma\sqrt{\lambda}\cdot t + K_2), & \text{if } S>\bar{S} 
	\end{cases}
\end{equation*}
Finally, we determine the constants $K_1$, $K_2$ by the initial condition $S_0$ and have 
\begin{align*}
	K_1 &= \tanh^{-1}(S_0/\bar{S})\\
	K_2 &= \coth^{-1}(S_0/\bar{S})
\end{align*}

>>>>

\subsection{Settings}

% Effort, Output and Gap
We assume two players, $i$ and $j$, compete for a prize $\theta>0$ in a contest. 
Winner gets the prize and loser gets nothing. 
The contest starts at time zero. 
At every time $t\ge0$, the representative player $i$ chooses an effort level $q_{i,t}$ and burdens a quadratic cost $C_i(q_{i,t}) = c_i q_{i,t}^2/2$, with a lower $c_i$ corresponding to higher ability. 
The dynamic of player $i$'s output follows 
\begin{equation}
dx_{i,t} = q_{i,t}dt + \sigma_idW_{i,t}
\end{equation}
Here $W_{i,t}$ is a standard Brownian motion and $\sigma_i>0$ measures her production risk. 
Moreover, denoted by $y_t$ the \textit{output gap} of player $i$ and $j$ at time $t$.

\subsubsection{Kalman Filter.}

We assume that the contest is equipped with a submission system that allows participants to upload their algorithms at any time and receive immediate feedback. 
For simplicity, we further assume that agents submit their intermediate results whenever they make progress. 
This setup enables the contest organizers to monitor all players’ progress $x_{i,t}$ and $x_{j,t}$ in real time (whenever there is a submission). 
The true outputs, evaluated by the system, is only known by the game designer but not the two players.

% Submission
Let' suppose the submission events of the representative player $i$ occur at times $(t^i_1, t^i_2, ...)$ following an inhomogeneous Poisson process driven by the intensity function $\tau_i(t)$.

% Signal
After each submission, the contest designer emits a \textit{public} signal of the real output $x_{i,t_k}$. 
The signal is ambiguous and the game holder controls the ambiguity. 
The dynamic of signal is 
\begin{equation}\label{signal}
	\hat{x}_{i,k} = x_{i,t_k} + \frac{v_{i,k}}{\sqrt{\lambda}}, \quad k=1, 2, ...
\end{equation}
where $v_{i,k}$ follows standard normal distribution and is independent with $(W_{i,t})$ and $(W_{j,t})$, and the parameter $\lambda$ is set by the game holder to control the precision of signals. 
The larger the $\lambda$, the more accurate the signal would be.

% Bayesian Player
The information set of both players at time $t \ge 0$ is  $I_{t} := \{\hat{x}_{i,k}, \hat{x}_{j,k} : 0\le t_k \le t\}$. 
Both players estimate the unknown outputs $x_{i,t}$ and $x_{j,t}$ purely based on the information set $I_t$ and hidden actions $q_{i,t}$ and $q_{j,t}$. 
Let $\tilde{x}_{i,t} \equiv E(x_{i,t}|I_t)$ be the estimated output gap and $S_{i,t} \equiv E[(\tilde{x}_{i,t}-x_{i,t})^2|I_t]$ be the estimation variance. 
The conditional distribution $y_{t}|I_t\sim\mathcal{N}(\tilde{y}_{t},S_{t}|I_t)$ is fully captured by the mean $\tilde{y}_{t}$ and variance $S_{t}$.

The evolution of $\tilde{x}_{i,t}$ and $S_{i,t}$ is characterized by a continuous-discrete Kalman Filter (CD-KF, \citealt{barrau2017invariant}, \citealt{frogerais2012various}), with the measurement of each step $k=1, 2, ...$ consisting of two phases: in
(1) prediction phase during time interval $(t_{k-1}, t_{k})$, equations are derived from those of Kalman-Bucy filter \citep{Bensoussan1992Control} without considering the Kalman gain: 
\begin{align}
	d\tilde{x}_{i,t} &= q_{i,t}dt
	\label{filtered-x}\\
	dS_{i,t} &= \sigma^2dt
	\label{filtered-S}
\end{align} 
and in (2) updation phase $t=t_k$, equations are 
\begin{align}
	\tilde{x}_{i,k}^{+} &= \tilde{x}_{i,k}^{-} + \frac{\lambda S_{i,t_k}^{-}}{\lambda S_{i,t_k}^{-}+1}\left(\hat{x}_{i,k} - \tilde{x}_{i,t_k}^{-}\right)
	\label{updated-x}\\
	S_{i,t_k}^{+} &= \frac{S_{i,t_k}^{-}}{\lambda S_{i,t_k}^{-}+1} 
\end{align}
If $\lambda = 0$, we have $S_t = S_0+\sigma^2t$, i.e., the estimation variance is increasing in time linearly. 
If $\lambda > 0$, the estimation error decreases after each update of the public leaderboard, i.e., $S_{i,t_k}^+ < S_{i,t_k}^-$.

\subsubsection{Dynamic Contest.}

% Contest Deadline
Following \cite{ryvkin2022fight}, let's consider a dynamic contest of two players with a fixed deadline. 
Suppose the contest is terminated when time $t=T>0$. 
Since the steady state estimation variance $\bar{S}$ is fixed as displayed above, the state of the game is fully characterized by a tuple $(\tilde{y}_t, t)$. 
At any time $0\le t<T$, player $i$ optimizes her effort level $q_{i,\tau}$ in the remaining contest period $\tau\in[t, T)$ according to the following optimization problem, 
\begin{equation}\label{v-def}
	V^i(\tilde{y}_{t}, t ; q_{j,t},\Theta_i) = \max_{\{q_{i,\tau}\}^T_{\tau=t}} 
	\mathbb{E}\left( \theta\cdot1_{\tilde{y}_T>0} - \int^T_tC_i(q_{i,\tau})d\tau \bigg|I_t\right) 
\end{equation}
where $\Theta_i \equiv\{\theta, \lambda, \sigma, c_i\}$, subject to constraints (\ref{filtered-x}), (\ref{filtered-S}) and $q_{i,\tau}\ge0$ for all $\tau\in[t,T)$.
The optimization problem for player $j$ is just symmetric to that of player $i$ as $V^j(\tilde{y}_t, t) = V^i(-\tilde{y}_t, t)$. 
The corresponding Hamilton-Jacobi-Bellman (HJB) equation for player $i$ is 
\begin{equation*}
0 = \max_{q_{i,t}\ge0}\left[-\frac{c_iq_i^2}{2} + V^i_{y}\cdot\left(q_{i,t}-q_{j,t}\right)+V^i_t + \frac{V^i_{yy}}{2}\lambda \bar{S}^2\right]
\end{equation*}
By definition, we have $\lambda \bar{S}^2 = \sigma^2$. 
Under the assumption of inner solution, we plug into the first order conditions $q_{i,t} = V^i_y/c_i$ and $q_{j,t} = -V^j_y/c_j$, we have the system of equations
\begin{equation*}
\begin{aligned}
\frac{1}{2c_i}(V^i_y)^2 + \frac{1}{c_j}V^i_yV^j_y + V^i_t + V^i_{yy}\frac{\sigma^2}{2} = 0\\
\frac{1}{2c_j}(V^j_y)^2 + \frac{1}{c_i}V^j_yV^i_y + V^j_t + V^j_{yy}\frac{\sigma^2}{2} = 0
\end{aligned}
\end{equation*}
subject to boundary conditions $V^i(-\infty, t) = 0$, $V^i(+\infty, t) = \theta$, $V^j(-\infty, t) = \theta$, $V^j(+\infty, t)=0$, $V^i(\tilde{y}_T, T) = \theta \cdot 1_{\tilde{y}_T > 0}$ and $V^j(\tilde{y}_T, T) = \theta \cdot 1_{\tilde{y}_T < 0}$.

@@

The likelihood function of any realization of this point process $\{\hat{t}^i_k\}_{k=1}^{N_i}$ is given by 
\begin{equation}\label{eq-ihpp-prob}
p\left(\{\hat{t}^i_k\}_{k=1}^{N_i} | \tau_i\right) = \exp\left\{-\int_{s\in\mathcal{T}}\tau_i(s)ds\right\}\prod_{k=1}^{N_i}\tau_i(\hat{t}^i_k)
\end{equation}

@@

Based on the following assumptions, we will next generate the submission data for two players in a data analysis competition: 
\begin{itemize}
	\item Total innovation uncertainty of the two players per day: $\sigma = 5$
	\item Capacities: $c_i = 1.5$ and $c_j = 2$
	\item Contest duration: from 2025-01-01 to 2025-03-31
	\item Starting point $\tilde{y}_0=0$
\end{itemize}

A central challenge in generating synthetic data lies in dynamically constructing a point process that conforms to an inhomogeneous Poisson process.
Inspired by the thinning technique \citep{lewis1979simulation}, we first generate candidate events according to a homogeneous Poisson process within each discrete time interval, using a fixed intensity $\tau_i^\star \ge \sup_t \tau_i(t)$, and distribute them uniformly across the interval.
We then apply the thinning procedure to determine whether each candidate event is accepted.

\begin{algorithm}
\caption{Synthetic Data Simulation}
\begin{algorithmic}
\STATE \textbf{Input:} 
	$c_i$, $c_j$, $\sigma$, $\lambda$, $r$, $\tau^\star$, $y_0$, $\tilde{y}_0$
\STATE Sample $\{s_k^{i(j)}\}^{N_{i(j)}}_{k=1}$ from homogeneous Poisson process $(\tau^\star_{i(j)})$ on $[0, T)$
\STATE Sample $\{u_k^{i(j)}\}^{N_{i(j)}}_{k=1}$ from uniform distribution on $[0, 1]$
\STATE Initialize $\ell=0$, $\ell^{i(j)}=0$, $\hat{y}_0 = 0$
\FOR{$t = 0$ to $T$}
    \STATE $m_{i(j)}(\tilde{y}_t, t)$  $\gets$ (\ref{eq-equilibrium-effort}), $\tau^{i(j)}(t)$ $\gets$ (\ref{eq-model-intensity})
    \STATE $\tilde{y}_{t+\Delta t}$ $\gets$ (\ref{eq-fintered-y-update})
    \FOR{ each $s_{k}^{i(j)} \subset [t, t+\Delta t)$} 
        \IF {$u_k^{i(j)} > \tau_{i(j)}(s_{k}^{i(j)}) / \tau^\star_{i(j)}$}
        	    \STATE $\hat{t}^{i(j)}_{\ell} \gets s_{k}^{i(j)}$; $\ell^{i(j)} \gets \ell^{i(j)}+1$
	    \hfill \COMMENT{Accept the submission event}
	    \STATE $\hat{y}_{\ell} \gets $; $\ell \gets \ell + 1$
	    \hfill \COMMENT{Update public leaderboard}
        \ENDIF
    \ENDFOR
\ENDFOR
\STATE \textbf{Output:} $(y_t)$, $(\tilde{y}_t)$, $\{\hat{t}_k^{i(j)}\}^{N_{i(j)}}_{k=1}$, $\{\hat{y}_k\}_{k=1}^{N_i + N_j}$
\end{algorithmic}
\end{algorithm}

@@

@@
